A Toolbox for Manipulating and Assessing Color Palettes

### 2019-12-02 pkg/colorspace/paper/Toolbox/article.Rmd
mention new adjust_transparency() in toolbox article
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/colorspace@522 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

>>>>

## Adjust transparency of colors

Alpha transparency is useful for making colors semi-transparent, e.g., for
overlaying different elements in graphics [@color:Wiki+Webcolors]. An alpha value
(or alpha channel) of 0 (or 00 in hex strings) corresponds to fully transparent and
an alpha value of 1 (or FF in hex strings) corresponds to fully opaque. If a color
hex string in \proglang{R} does not provide an explicit alpha transparency, the
color is assumed to be fully opaque.

The `adjust_transparency()` function can be used to adjust the alpha transparency
of a set of colors. It always returns a hex color specification. This hex color
can have the alpha transparency added/removed/modified depending on the
specification of the argument `alpha`:

* `alpha = NULL`: Returns a hex vector with alpha transparency only if needed.
  Thus, it keeps the alpha transparency for the colors (if any) but only if
  different from opaque.
* `alpha = TRUE`: Returns a hex vector with alpha transparency
  for all colors, using opaque (FF) as the default if missing.
* `alpha = FALSE`: Returns a hex vector without alpha transparency for
  all colors (even if the original colors had non-opaque alpha).
* `alpha` numeric: Returns a hex vector with alpha transparency for
  all colors set to the `alpha` argument (recycled if necessary).

For illustration, the transparency of a single black color is modified to three
alpha levels: fully transparent, semi-transparent, and fully opaque, respectively.
Black can be equivalently specified by name (`"black"`), hex string (`"#000000"`),
or integer position in the palette (`1`).

```{r transparency-black}
adjust_transparency("black",   alpha = c(0, 0.5, 1))
adjust_transparency("#000000", alpha = c(0, 0.5, 1))
adjust_transparency(1,         alpha = c(0, 0.5, 1))
```

Subsequently, different settings of `alpha` are illustrated for adjusting
a vector with three shades of gray, specified by name (`gray`, opaque),
opaque hex string (`"#BEBEBE"`), and semi-transparent hex string (`"#BEBEBE80"`).
Four types of adjustment are shown: only if necessary (`alpha = NULL`),
add (`alpha = TRUE`), remove (`alpha = FALSE`), or modify (`alpha = 0.8`).

```{r transparency-gray}
x <- c("gray", "#BEBEBE", "#BEBEBE80")
adjust_transparency(x, alpha = NULL)
adjust_transparency(x, alpha = TRUE)
adjust_transparency(x, alpha = FALSE)
adjust_transparency(x, alpha = 0.8)
```
